---
title: "Standard Errors and Confidence Intervals"
author: "Lauren K. Perez"
date: "1/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "your/file/path") #Either add your working directory here or comment this out
```

# Confidence Interval and Margin of Error for a Poll

Go to [Polling Report](www.pollingreport.com) and find a poll that looks interesting to you.  You should choose one where you can reasonably collapse the responses into two categories.  

First, using mathematical formulas, figure out:

  * the standard deviation
  * the standard error
  * the 95% confidence interval
  * the 99% confidence interval
  * the margin of error at 95% 
  * the margin of error at 99% 
  
Next, create a vector in R with the poll responses.  You can use the rep() function to create vectors of 1s and 0s, and then the append() function to combine them. Now check that the mean and standard deviation you found above are correct.

You can also use R to check the confidence interval.  Use the following code to calculate the 95% confidence interval.  First, create an object called `se` that is the standard error that you calculated above and `mean` that is the mean from above.  Then replace `vector` with whatever you named your vector of poll responses. 

```{r}
c(mean - qnorm(0.975)*se, mean + qnorm(0.975)*se)
```

This uses 1.96(SE) rather thane than 2(SE), so this confidence interval may be slightly smaller than the one you calculated above. 

Now use R to check the 99% confidence interval. 

# Observing the Central Limit Theorem on the College Scorecard Data

```{r}
library(haven)
data2<-read_dta("/Users/kevinxyliu/Desktop/CollegeScorecard1415_forR.dta")

# clean up the data
data2$opeid6[which(data2$opeid6=="NULL")]<-NA
data2$instnm[which(data2$instnm=="NULL")]<-NA
data2$city[which(data2$city=="NULL")]<-NA
data2$stabbr[which(data2$stabbr=="NULL")]<-NA
data2$accredagency[which(data2$accredagency=="NULL")]<-NA
data2$hcm2[which(data2$hcm2=="NULL")]<-NA
data2$main[which(data2$main=="NULL")]<-NA
data2$numbranch[which(data2$numbranch=="NULL")]<-NA
data2$preddeg[which(data2$preddeg=="NULL")]<-NA
data2$highdeg[which(data2$highdeg=="NULL")]<-NA
data2$control[which(data2$control=="NULL")]<-NA
data2$menonly[which(data2$menonly=="NULL")]<-NA
data2$womenonly[which(data2$womenonly=="NULL")]<-NA
data2$adm_rate[which(data2$adm_rate=="NULL")]<-NA
data2$relaffil[which(data2$relaffil=="NULL")]<-NA
data2$actcm25[which(data2$actcm25=="NULL")]<-NA
data2$actcm75[which(data2$actcm75=="NULL")]<-NA
data2$sat_avg[which(data2$sat_avg=="NULL")]<-NA
data2$ugds[which(data2$ugds=="NULL")]<-NA
data2$ugds_white[which(data2$ugds_white=="NULL")]<-NA
data2$ugds[which(data2$ugds=="NULL")]<-NA
data2$ugds_black[which(data2$ugds_black=="NULL")]<-NA
data2$ugds_hisp[which(data2$ugds_hisp=="NULL")]<-NA
data2$ugds_asian[which(data2$ugds_asian=="NULL")]<-NA
data2$npt4_pub[which(data2$npt4_pub=="NULL")]<-NA
data2$npt4_priv[which(data2$npt4_priv=="NULL")]<-NA
data2$costt4_a[which(data2$costt4_a=="NULL")]<-NA
data2$tuitionfee_in[which(data2$tuitionfee_in=="NULL")]<-NA
data2$tuitionfee_out[which(data2$tuitionfee_out=="NULL")]<-NA
data2$inexpfte[which(data2$inexpfte=="NULL")]<-NA
data2$avgfacsal[which(data2$avgfacsal=="NULL")]<-NA
data2$pctpell[which(data2$pctpell=="NULL")]<-NA
data2$c200_4[which(data2$c200_4=="NULL")]<-NA
data2$c200_l4[which(data2$c200_l4=="NULL")]<-NA
data2$ret_ft4[which(data2$ret_ft4=="NULL")]<-NA
data2$ret_ftl4[which(data2$ret_ftl4=="NULL")]<-NA
data2$ret_pt4[which(data2$ret_pt4=="NULL")]<-NA
data2$ret_ptl4[which(data2$ret_ptl4=="NULL")]<-NA
data2$pctfloan[which(data2$pctfloan=="NULL")]<-NA
data2$ug25abv[which(data2$ug25abv=="NULL")]<-NA
data2$cdr3[which(data2$cdr3=="NULL")]<-NA
data2$rpy_1yr_rt[which(data2$rpy_1yr_rt=="NULL")]<-NA
data2$female_rpy_1yr_rt[which(data2$female_rpy_1yr_rt=="NULL")]<-NA
data2$male_rpy_1yr_rt[which(data2$male_rpy_1yr_rt=="NULL")]<-NA
data2$rpy_3yr_rt[which(data2$rpy_3yr_rt=="NULL")]<-NA
data2$female_rpy_3yr_rt[which(data2$female_rpy_3yr_rt=="NULL")]<-NA
data2$male_rpy_3yr_rt[which(data2$male_rpy_3yr_rt=="NULL")]<-NA
data2$rpy_5yr_rt[which(data2$rpy_5yr_rt=="NULL")]<-NA
data2$female_rpy_5yr_rt[which(data2$female_rpy_5yr_rt=="NULL")]<-NA
data2$male_rpy_5yr_rt[which(data2$male_rpy_5yr_rt=="NULL")]<-NA
data2$inc_pct_lo[which(data2$inc_pct_lo=="NULL")]<-NA
data2$par_ed_pct_1stgen[which(data2$par_ed_pct_1stgen=="NULL")]<-NA
data2$inc_pct_m1[which(data2$inc_pct_m1=="NULL")]<-NA
data2$inc_pct_m2[which(data2$inc_pct_m2=="NULL")]<-NA
data2$inc_pct_h1[which(data2$inc_pct_h1=="NULL")]<-NA
data2$inc_pct_h2[which(data2$inc_pct_h2=="NULL")]<-NA
data2$debt_mdn[which(data2$debt_mdn=="NULL")]<-NA
data2$grad_debt_mdn[which(data2$grad_debt_mdn=="NULL")]<-NA
data2$wdraw_debt_mdn[which(data2$wdraw_debt_mdn=="NULL")]<-NA
data2$faminc[which(data2$faminc=="NULL")]<-NA
data2$md_faminc[which(data2$md_faminc=="NULL")]<-NA
data2$faminc_ind[which(data2$faminc_ind=="NULL")]<-NA
data2$mn_earn_wne_p10[which(data2$mn_earn_wne_p10=="NULL")]<-NA
data2$md_earn_wne_p10[which(data2$md_earn_wne_p10=="NULL")]<-NA
data2$gt_25k_p10[which(data2$gt_25k_p10=="NULL")]<-NA
data2$c100_4[which(data2$c100_4=="NULL")]<-NA
data2$c100_l4[which(data2$c100_l4=="NULL")]<-NA
data2$ugds_men[which(data2$ugds_men=="NULL")]<-NA
data2$ugds_women[which(data2$ugds_women=="NULL")]<-NA
data2$opeid6[which(data2$opeid6=="PrivacySuppressed")]<-NA
data2$instnm[which(data2$instnm=="PrivacySuppressed")]<-NA
data2$city[which(data2$city=="PrivacySuppressed")]<-NA
data2$stabbr[which(data2$stabbr=="PrivacySuppressed")]<-NA
data2$accredagency[which(data2$accredagency=="PrivacySuppressed")]<-NA
data2$hcm2[which(data2$hcm2=="PrivacySuppressedL")]<-NA
data2$main[which(data2$main=="PrivacySuppressed")]<-NA
data2$numbranch[which(data2$numbranch=="PrivacySuppressed")]<-NA
data2$preddeg[which(data2$preddeg=="PrivacySuppressed")]<-NA
data2$highdeg[which(data2$highdeg=="PrivacySuppressed")]<-NA
data2$control[which(data2$control=="PrivacySuppressed")]<-NA
data2$menonly[which(data2$menonly=="PrivacySuppressed")]<-NA
data2$womenonly[which(data2$womenonly=="PrivacySuppressed")]<-NA
data2$adm_rate[which(data2$adm_rate=="PrivacySuppressed")]<-NA
data2$relaffil[which(data2$relaffil=="PrivacySuppressed")]<-NA
data2$actcm25[which(data2$actcm25=="PrivacySuppressed")]<-NA
data2$actcm75[which(data2$actcm75=="PrivacySuppressed")]<-NA
data2$sat_avg[which(data2$sat_avg=="PrivacySuppressed")]<-NA
data2$ugds[which(data2$ugds=="PrivacySuppressed")]<-NA
data2$ugds_white[which(data2$ugds_white=="PrivacySuppressed")]<-NA
data2$ugds[which(data2$ugds=="PrivacySuppressed")]<-NA
data2$ugds_black[which(data2$ugds_black=="PrivacySuppressed")]<-NA
data2$ugds_hisp[which(data2$ugds_hisp=="PrivacySuppressed")]<-NA
data2$ugds_asian[which(data2$ugds_asian=="PrivacySuppressed")]<-NA
data2$npt4_pub[which(data2$npt4_pub=="PrivacySuppressed")]<-NA
data2$npt4_priv[which(data2$npt4_priv=="PrivacySuppressed")]<-NA
data2$costt4_a[which(data2$costt4_a=="PrivacySuppressed")]<-NA
data2$tuitionfee_in[which(data2$tuitionfee_in=="PrivacySuppressed")]<-NA
data2$tuitionfee_out[which(data2$tuitionfee_out=="PrivacySuppressed")]<-NA
data2$inexpfte[which(data2$inexpfte=="PrivacySuppressed")]<-NA
data2$avgfacsal[which(data2$avgfacsal=="PrivacySuppressed")]<-NA
data2$pctpell[which(data2$pctpell=="PrivacySuppressed")]<-NA
data2$c200_4[which(data2$c200_4=="PrivacySuppressed")]<-NA
data2$c200_l4[which(data2$c200_l4=="PrivacySuppressed")]<-NA
data2$ret_ft4[which(data2$ret_ft4=="PrivacySuppressed")]<-NA
data2$ret_ftl4[which(data2$ret_ftl4=="PrivacySuppressed")]<-NA
data2$ret_pt4[which(data2$ret_pt4=="PrivacySuppressed")]<-NA
data2$ret_ptl4[which(data2$ret_ptl4=="PrivacySuppressed")]<-NA
data2$pctfloan[which(data2$pctfloan=="PrivacySuppressed")]<-NA
data2$ug25abv[which(data2$ug25abv=="PrivacySuppressed")]<-NA
data2$cdr3[which(data2$cdr3=="PrivacySuppressed")]<-NA
data2$rpy_1yr_rt[which(data2$rpy_1yr_rt=="PrivacySuppressed")]<-NA
data2$female_rpy_1yr_rt[which(data2$female_rpy_1yr_rt=="PrivacySuppressed")]<-NA
data2$male_rpy_1yr_rt[which(data2$male_rpy_1yr_rt=="PrivacySuppressed")]<-NA
data2$rpy_3yr_rt[which(data2$rpy_3yr_rt=="PrivacySuppressed")]<-NA
data2$female_rpy_3yr_rt[which(data2$female_rpy_3yr_rt=="PrivacySuppressed")]<-NA
data2$male_rpy_3yr_rt[which(data2$male_rpy_3yr_rt=="PrivacySuppressed")]<-NA
data2$rpy_5yr_rt[which(data2$rpy_5yr_rt=="PrivacySuppressed")]<-NA
data2$female_rpy_5yr_rt[which(data2$female_rpy_5yr_rt=="PrivacySuppressed")]<-NA
data2$male_rpy_5yr_rt[which(data2$male_rpy_5yr_rt=="PrivacySuppressed")]<-NA
data2$inc_pct_lo[which(data2$inc_pct_lo=="PrivacySuppressed")]<-NA
data2$par_ed_pct_1stgen[which(data2$par_ed_pct_1stgen=="PrivacySuppressed")]<-NA
data2$inc_pct_m1[which(data2$inc_pct_m1=="PrivacySuppressed")]<-NA
data2$inc_pct_m2[which(data2$inc_pct_m2=="PrivacySuppressed")]<-NA
data2$inc_pct_h1[which(data2$inc_pct_h1=="PrivacySuppressed")]<-NA
data2$inc_pct_h2[which(data2$inc_pct_h2=="PrivacySuppressed")]<-NA
data2$debt_mdn[which(data2$debt_mdn=="PrivacySuppressed")]<-NA
data2$grad_debt_mdn[which(data2$grad_debt_mdn=="PrivacySuppressed")]<-NA
data2$wdraw_debt_mdn[which(data2$wdraw_debt_mdn=="PrivacySuppressed")]<-NA
data2$faminc[which(data2$faminc=="PrivacySuppressed")]<-NA
data2$md_faminc[which(data2$md_faminc=="PrivacySuppressed")]<-NA
data2$faminc_ind[which(data2$faminc_ind=="PrivacySuppressed")]<-NA
data2$mn_earn_wne_p10[which(data2$mn_earn_wne_p10=="PrivacySuppressed")]<-NA
data2$md_earn_wne_p10[which(data2$md_earn_wne_p10=="PrivacySuppressed")]<-NA
data2$gt_25k_p10[which(data2$gt_25k_p10=="PrivacySuppressed")]<-NA
data2$c100_4[which(data2$c100_4=="PrivacySuppressed")]<-NA
data2$c100_l4[which(data2$c100_l4=="PrivacySuppressed")]<-NA
data2$ugds_men[which(data2$ugds_men=="PrivacySuppressed")]<-NA
data2$ugds_women[which(data2$ugds_women=="PrivacySuppressed")]<-NA
data2$relaffil[which(data2$relaffil=="-2")]<-NA
data2$preddeg[which(data2$preddeg=="0")]<-NA

# convert data into numerics
data2$opeid6<-as.numeric(data2$opeid6)
data2$hcm2<-as.numeric(data2$hcm2)
data2$main<-as.numeric(data2$main)
data2$numbranch<-as.numeric(data2$numbranch)
data2$preddeg<-as.numeric(data2$preddeg)
data2$highdeg<-as.numeric(data2$highdeg)
data2$control<-as.numeric(data2$control)
data2$menonly<-as.numeric(data2$menonly)
data2$womenonly<-as.numeric(data2$womenonly)
data2$relaffil<-as.numeric(data2$relaffil)
data2$adm_rate<-as.numeric(data2$adm_rate)
data2$actcm25<-as.numeric(data2$actcm25)
data2$actcm75<-as.numeric(data2$actcm75)
data2$sat_avg<-as.numeric(data2$sat_avg)
data2$ugds<-as.numeric(data2$ugds)
data2$ugds_white<-as.numeric(data2$ugds_white)
data2$ugds_black<-as.numeric(data2$ugds_black)
data2$ugds_hisp<-as.numeric(data2$ugds_hisp)
data2$ugds_asian<-as.numeric(data2$ugds_asian)
data2$npt4_pub<-as.numeric(data2$npt4_pub)
data2$npt4_priv<-as.numeric(data2$npt4_priv)
data2$costt4_a<-as.numeric(data2$costt4_a)
data2$tuitionfee_in<-as.numeric(data2$tuitionfee_in)
data2$tuitionfee_out<-as.numeric(data2$tuitionfee_out)
data2$tuitfte<-as.numeric(data2$tuitfte)
data2$inexpfte<-as.numeric(data2$inexpfte)
data2$avgfacsal<-as.numeric(data2$avgfacsal)
data2$pctpell<-as.numeric(data2$pctpell)
data2$c200_4<-as.numeric(data2$c200_4)
data2$c200_l4<-as.numeric(data2$c200_l4)
data2$ret_ft4<-as.numeric(data2$ret_ft4)
data2$ret_ftl4<-as.numeric(data2$ret_ftl4)
data2$ret_pt4<-as.numeric(data2$ret_pt4)
data2$ret_ptl4<-as.numeric(data2$ret_ptl4)
data2$pctfloan<-as.numeric(data2$pctfloan)
data2$ug25abv<-as.numeric(data2$ug25abv)
data2$cdr3<-as.numeric(data2$cdr3)
data2$rpy_1yr_rt<-as.numeric(data2$rpy_1yr_rt)
data2$female_rpy_1yr_rt<-as.numeric(data2$female_rpy_1yr_rt)
data2$male_rpy_1yr_rt<-as.numeric(data2$male_rpy_1yr_rt)
data2$rpy_3yr_rt<-as.numeric(data2$rpy_3yr_rt)
data2$female_rpy_3yr_rt<-as.numeric(data2$female_rpy_3yr_rt)
data2$male_rpy_3yr_rt<-as.numeric(data2$male_rpy_3yr_rt)
data2$rpy_5yr_rt<-as.numeric(data2$rpy_5yr_rt)
data2$female_rpy_5yr_rt<-as.numeric(data2$female_rpy_5yr_rt)
data2$male_rpy_5yr_rt<-as.numeric(data2$male_rpy_5yr_rt)
data2$inc_pct_lo<-as.numeric(data2$inc_pct_lo)
data2$par_ed_pct_1stgen<-as.numeric(data2$par_ed_pct_1stgen)
data2$inc_pct_m1<-as.numeric(data2$inc_pct_m1)
data2$inc_pct_m2<-as.numeric(data2$inc_pct_m2)
data2$inc_pct_h1<-as.numeric(data2$inc_pct_h1)
data2$inc_pct_h2<-as.numeric(data2$inc_pct_h2)
data2$debt_mdn<-as.numeric(data2$debt_mdn)
data2$grad_debt_mdn<-as.numeric(data2$grad_debt_mdn)
data2$wdraw_debt_mdn<-as.numeric(data2$wdraw_debt_mdn)
data2$faminc<-as.numeric(data2$faminc)
data2$md_faminc<-as.numeric(data2$md_faminc)
data2$faminc_ind<-as.numeric(data2$faminc_ind)
data2$mn_earn_wne_p10<-as.numeric(data2$mn_earn_wne_p10)
data2$md_earn_wne_p10<-as.numeric(data2$md_earn_wne_p10)
data2$gt_25k_p10<-as.numeric(data2$gt_25k_p10)
data2$c100_4<-as.numeric(data2$c100_4)
data2$c100_l4<-as.numeric(data2$c100_l4)
data2$ugds_men<-as.numeric(data2$ugds_men)
data2$ugds_women<-as.numeric(data2$ugds_women)
```

Begin by opening the College Scorecard data and doing any necessary cleaning.  The variables you will need for today are `adm_rate`, the `control` variable and a dummy variable for whether or not a school is selective (if it has an admission rate below .35). Remember that `contro`l is coded 1 for public, 2 for private non-profit, and 3 for private for-profit. 

```{r}
data2$selective <-0
data2$selective[which(data2$adm_rate < 0.3)]<-1
data2$selective[which(is.na(data2$adm_rate))]<-NA

```

Let's begin by testing out the Central Limit Theorem using the `adm_rate` variable.  First, we need to get just the non-missing data, which we can do using the line of code below. (If we draw samples from the data including missing data, our standard error will be larger than expected.)

```{r}
nonNA_adm_rate <- data2$adm_rate[which(!is.na(data2$adm_rate))]
```

Next, you should find the population mean ($\mu$) and the population standard deviation ($\sigma$) of the admission rate data.  Then, produce a histogram of the data.  Is the underlying data normally distributed?
```{r}
mean(nonNA_adm_rate)
sd(nonNA_adm_rate)
```

Next, set up an empty vector to store the sample means in and a loop that will draw 10,000 samples of 100 colleges from the data, storing their means in your vector. Remember to draw the samples from the `nonNA_adm_rate` vector, rather than from `data$adm_rate`. 
```{r}
sample_mean<-rep(NA,10000)
for (f in 1:10000){
  sam<-sample(nonNA_adm_rate,100)
  sample_mean[f]<-mean(sam)
}
mean(sample_mean)
sd(sample_mean)
```

Using your new vector of sample means, find the mean and the standard deviation.  Are these what the central limit theorem would lead you to expect?  (Remember, the standard deviation of the sampling distribution is the standard error of the underlying data.) 

Make a histogram of the sample means.  Are they normally distributed around the population mean?

```{r}
hist(sample_mean)
```

#Graphing Exercise

During the second week, we learned how to do bar graphs in R.  Today let's learn how to do bar graphs for multiple populations. 

Start by splitting the data into two datasets - one for selective schools and one for non-selective schools. 

```{r}
CSselective <- data2[data2$selective==1,]

CSnonselective <- data2[data2$selective==0,]
```


Compare how these two groups differ across the `control` variable and the `adm_rate` variable using descriptive statistics.
```{r}
table(CSselective$control)
table(CSnonselective$control)
```

Next, let's try making a bar graph that plots selective and non-selective schools frequencies for the control variable next to one another.  You will need to adjust the data and variable names below to match yours.

```{r}
tabs <- table(CSselective$control)
tabns <- table(CSnonselective$control)
controlsns <- rbind(tabs, tabns)

plot <- barplot(controlsns, beside = TRUE,
        names.arg=c("Public", "Private Non-Profit", "Private For-Profit"),
        cex.names = 0.8,
        main = "Type of School Control by Selectivity", 
        legend = c("Selective", "Non-Selective")
        )
```





